---
title: "Aiptasia Manual Zoox Counts"
output:
  pdf_document:
    fig_caption: true
  html_document:
    df_print: paged
  md_document:
    variant: gfm
    df_print: paged
  word_document: default
classoption: portrait
editor_options: 
  chunk_output_type: inline
---
```{r global_options, echo=TRUE}
rm(list = ls())
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, results='asis')
```

```{r}
library("knitr")
library("kableExtra")
library("tidyr")
library("dplyr")
library("knitr")
library("pander")
```

Create a dataframe of the weights of the anemones in milligrams and the counts of zoox in 8 squares of hemocytometer.

```{r}
#Weight of anemones in milligrams
mgWeight <- c(103, 139, 156, 182, 361, 67, 126, 85, 68, 65, 142)
#Counts of zoox in 8 squares of hemocytometer.
Count <- c(187, 132, 107, 71, 111, 1,  2,  5, 43,  2,  8)
#Treatment groups
Menthol <- c("Control", "Control", "Control", "Control", "Control", 
             "Menthol","Menthol","Menthol", "Control", "Menthol", "Menthol")
dfCounts <- data.frame(Weight=mgWeight, Count=Count, Menthol=Menthol)
```

Calculate zoox per anemone. Each anemone homogenized in 1000 uL filtered seawater. Hemocytometer volume is 100 uL.

```{r}
#Small squares counted
squaresCounted <- 8
#Volume of small square in mL
volumeSmallSquare <- 0.000004 
volumeSquaresCounted <- (volumeSmallSquare) * squaresCounted
```
```{r}
#Calculate zoox/mL
dfCounts$perMl <- with(dfCounts, (Count/volumeSquaresCounted))
#Log-transform the zoox counts
dfCounts$ZooxPerMgLog <- with(dfCounts, log2(perMl/Weight))
#Calculate zoox/weight/volume
dfCounts$ZooxPerMg <- with(dfCounts, perMl/Weight)
```

\newpage

```{r}
dfCounts <- subset(dfCounts, select=c("Menthol","Count", "perMl","Weight", "ZooxPerMg", "ZooxPerMgLog"))
kable(dfCounts, caption="Symbiodinium Count Summary", digits=2)
library(plotrix)
mentholCounts <- subset(dfCounts, Menthol == 'Menthol')
controlCounts <- subset(dfCounts, Menthol == 'Control')
mean(mentholCounts$ZooxPerMg)
std.error(mentholCounts$ZooxPerMg)
mean(controlCounts$ZooxPerMg)
std.error(controlCounts$ZooxPerMg)
```

\newpage
Shapiro Test Zoox/Mg

```{r}
shapiro.test(dfCounts$ZooxPerMg)
```

\newpage

```{r}
qqnorm(dfCounts$ZooxPerMg)
```

\newpage

```{r}
shapiro.test(dfCounts$ZooxPerMgLog)
```

```{r}
qqnorm(dfCounts$ZooxPerMgLog)
```


\newpage

One-way ANOVA of the zoox counts

```{r}
aov.zoox.log <- aov(ZooxPerMg ~ Menthol, data=dfCounts)
summary(aov.zoox.log)
meansTable <- model.tables(aov.zoox.log, "means", se.contrast=T)
se.contrast(aov.zoox.log, list(Menthol == "Menthol", Menthol == "Control"))
```
Boxplot of raw counts

```{r fig.cap="Zoox Count"}
boxplot(ZooxPerMg ~ Menthol, data=dfCounts)
```

One-way ANOVA of the log-transformed zoox counts

```{r}
aov.zoox.log <- aov(ZooxPerMgLog ~ Menthol, data=dfCounts)
summary(aov.zoox.log)
meansTable <- model.tables(aov.zoox.log, "means", se=F)
stdErr <- se.contrast(aov.zoox.log, list(Menthol == "Menthol", Menthol == "Control"))
#means<- unclass(meansTable)
mentholMean <- meansTable$tables$Menthol[1]
controlMean <- meansTable$tables$Menthol[2]
alogStdErr <- 2^stdErr
alogMentholMean <- 2^mentholMean
alogControlMean <- 2^controlMean
foldChange <- 2^(mentholMean - controlMean)
alogStdErr
alogMentholMean
alogControlMean
foldChange
```

\newpage
Boxplot of log-transformed counts

```{r fig.cap="Zoox Count"}
boxplot(ZooxPerMgLog ~ Menthol, data=dfCounts)
```

