---
title: "ZooxCount"
output: 
  pdf_document: 
      fig_caption: true
  html_document:
    df_print: paged
  md_document:
    variant: gfm
    df_print: paged
  word_document: default
---

**Measuring Bleaching**

To measure the degree of bleaching after menthol exposure, 6 bleached and 6 unbleached anemones were randomly selected for zoox counting at the beginning of the experiment. After the bleaching cycle the selected anemones were weighed and homogenized in filtered seawater. One homogenate tube was dropped and the homogenate lost, leaving 6 control and 5 menthol samples for counting. The homogenate was viewed using a hemocytometer. Images were captured and zoox were counted using the BioConductor package EBImage. In addition to counting zoox with EBImage, zoox were manually counted using ImageJ. The images showing manual counts are available as supplemental data in the ManualCountImageJ folder. The automated method tended to count circular chunks of cellular debris as zoox, even though by visual inspection the chunks were clearly not zoox. Despite the over-counting of non-zoox cellular debris, both the automated and manual counts showed a significant reduction in zoox after treatment with menthol. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, results='asis')
## try http:// if https:// URLs are not supported
library("EBImage")
library("knitr")
```

```{r}
#Read in all hemocytometer images
dir <- 'ZooxCount/'
files <- list.files(dir, pattern="Aip*")
#Output directory for images showing counted zoox
outDir <- 'ZooxCounted/'
#files
```

```{r}
#Initialize a data frame to store counts.
dfCounts <- data.frame()

for (ifile in files){
  x <- readImage(paste0(dir, ifile))
  #Output the original image
  #display(x, method="raster")
  #Convert the image to grayscale
  colorMode(x) <- "Grayscale"
  #Keep just the frame with the grayscale image
  x <- getFrame(x, 3)
  #Find disc-shaped areas about the size of zoox
  disc = makeBrush(23, "disc")
  disc = disc / sum(disc)
  offset = 0.05
  x = filter2( x, disc )
  #Convert to black and white for a binary determination - zoox or not?
  x <- bwlabel(x < .48)
  #Output the counted image to show which regions were counted as zoox
  #display(x, method="raster")
  #Write the modified image as a separate file
  writeImage(x, file=paste0(outDir, ifile))
  #Get the count and add to the data frame
  count <- max(x)
  #print(count)
  dfCounts[ifile, "Count"] <- count
}
```

```{r}
Menthol <- c("Control", "Control", "Control", "Control", "Control", 
             "Menthol","Menthol","Menthol", "Control", "Menthol", "Menthol")
#Add the weights in milligrams to the data frame
mgWeight <- c(103, 139, 156, 182, 361, 67, 126, 85, 68, 65, 142)
#Number of squares in hemocytometer
hemoGrid <- 15*15
#Squares counted
squaresCounted <- 8
#Hemocytometer volume
volumeMl <- .1 
#Add manual counts of zoox.
dfCounts$ManualCount <- c(187, 132, 107, 71, 111, 1,  2,  5, 43,  2,  8)
#Calculate zoox/mL
dfCounts$perMl <- with(dfCounts, (Count/volumeMl) * (squaresCounted /hemoGrid))
dfCounts$perMlManual <- with(dfCounts, (ManualCount/volumeMl) * (squaresCounted /hemoGrid))
#Add menthol information to data frame
dfCounts$Menthol <- Menthol
#Add weight information to data frame
dfCounts$Weight <- mgWeight
#Log-transform the zoox counts
dfCounts$logCount <- with(dfCounts, log(perMl))
dfCounts$logCountManual <- with(dfCounts, log(perMlManual))
#Calculate zoox/weight/volume
dfCounts$ZooxPerMgMl <- with(dfCounts, perMl/Weight)
dfCounts$ZooxPerMgMlManual <- with(dfCounts, perMlManual/Weight)
#Calculate log zoox/weight/volume
dfCounts$ZooxPerMgMlLog <- with(dfCounts, logCount/Weight)
dfCounts$ZooxPerMgMlLogManual <- with(dfCounts, logCountManual/Weight)
```

\newpage

One-way ANOVA of the log-transformed zoox counts showed a significant difference in the zoox count between bleached and un-bleached anemones.

```{r}
aov.zoox.log <- aov(ZooxPerMgMlLog ~ Menthol, data=dfCounts)

zooxSumm <- as.data.frame(unclass(summary(aov.zoox.log)))
kable(zooxSumm, caption = "One-way ANOVA, automated zoox counts")
```

\newpage

Reduction in zoox count for bleached anemones.

```{r fig.cap="Automated Zoox Count"}
boxplot(ZooxPerMgMl ~ Menthol, data=dfCounts)
```

\newpage

One-way ANOVA of the log-transformed zoox manual counts showed a significant difference in the zoox count between bleached and un-bleached anemones.

```{r}
aov.zoox.logManual <- aov(ZooxPerMgMlLogManual ~ Menthol, data=dfCounts)
zooxSummManual <- as.data.frame(unclass(summary(aov.zoox.logManual)))
kable(zooxSummManual, caption = "One-way ANOVA, manual zoox counts")
```

\newpage

Reduction in zoox count for bleached anemones.

```{r fig.cap="Manual Zoox Count"}
boxplot(ZooxPerMgMl ~ Menthol, data=dfCounts)
```

\newpage

One-way ANOVA of the anemone weights did not show a significant difference in weight between the bleached and un-bleached anemones. 

```{r}
aov.weight <- aov(Weight ~ Menthol, data=dfCounts)
weightSumm <- as.data.frame(unclass(summary(aov.weight)))
kable(weightSumm, caption = "One-way ANOVA, anemone weight")
```

\newpage

```{r fig.cap="Weights of bleached and un-bleached anemones"}
boxplot(Weight ~ Menthol, data=dfCounts)
```

\newpage

```{r fig.cap="Symbiodinium Counts Bleached and Un-Bleached"}
dfCounts <- subset(dfCounts, select=c("Menthol","Count","ManualCount", "perMl","logCount", "Weight", "ZooxPerMgMl"))
kable(dfCounts)
```

